taxadb: A High-Performance Local Taxonomic Database Interface

### 2019-12-12 paper/manuscript.Rmd
Clean up

<!--
`taxadb` does not require the use of an external database system, but will automatically make use of the fastest database server available. `taxadb` can also use any DBI compliant database accessible from R, such as MySQL or Postgres [see @DBI], though the default `duckdb` or `MonetDBLite` will be both significantly faster and easier to use.  
-->

<!--  Next 2 paragraphs should possibly be left to a manual, but are likely to be overlooked aspects there... -->
By default `taxadb` creates a MonetDBLite database instance [@monetdblite], a columnar-oriented relational database requiring no additional installation while also providing persistent disk-based storage. `taxadb` will automatically handle opening, caching, and closing the database connection.

<!-- 
now sidestepped

but note that one limitation of MonetDBLite is the restriction against concurrent access: two separate R sessions on the same machine cannot both access the database at the same time. Users who need concurrent access from multiple sessions may wish to use a standalone MonetDB server or other database server instead.
-->

@@

`get_ids` offers a first pass at matching scientifc names to id, but names may remain unresolved for a number of reasons. First, a name may match to multiple accepted names, as in the case of a species that has been split. By design, these cases are left to be resolved by the researcher using the 'filter_' functions to filter underlying taxonomic tables for additional information. A name may also be unresolved due to typos or improper formatting. `clean_names` addresses common formatting issues such as the inclusion of missing species epithets (e.g. `Accipiter sp.`) that prevent matches to the Genus, or intraspecific epithets such as `Colaptes auratus cafer` that prevent matches to the bionmial name. These modifications are not appropriate in all settings and should be used with care (Language from vignette).

@@

<!--
Table version of below - citation, number of unique identifier

- the Integrated Taxonomic Information System [ITIS; @itis], original formed to standardize taxonomic name usage across many agencies in the United States federal government, 
- the National Center for Biological Information's (NCBI) Taxonomy database, [@ncbi]
- Catalogue of Life (COL) Annual Species list. (2018 version at present)
- Global Biodiversity Information Facility Taxonomic Backbone [GBIF; @gbif].  
- FishBase [@fishbase] taxonomic names
- SeaLifeBase [@sealifebase] taxonomic names
- Wikidata [@jorrit]
- Open Tree Taxonomy, OTT
- International Union for Conservation of Nature and Natural Resources (IUCN) Red list @iucn
-->

@@

* cite other tools in the space (taxize + citations)
* Frandz citations - grumpy systemitists; we're not doing anything inferential
* no spell check/name modification
* could be more than one name for rank
* versioning paragraph - argument for version like provider; we process every so often
  * comment on the release schedule, rolling is not always up to date

>>>>

By default `taxadb` creates a MonetDBLite database instance [@monetdblite], a columnar-oriented relational database requiring no additional installation while also providing persistent disk-based storage. `taxadb` will automatically handle opening, caching, and closing the database connection.

@@

`get_ids` offers a first pass at matching scientifc names to id, but names may remain unresolved for a number of reasons. First, a name may match to multiple accepted names, as in the case of a species that has been split. By design, these cases are left to be resolved by the researcher using the 'filter_' functions to filter underlying taxonomic tables for additional information. A name may also be unresolved due to typos or improper formatting. `clean_names` addresses common formatting issues such as the inclusion of missing species epithets (e.g. `Accipiter sp.`) that prevent matches to the Genus, or intraspecific epithets such as `Colaptes auratus cafer` that prevent matches to the bionmial name. These modifications are not appropriate in all settings and should be used with care (Language from vignette). Spell check of input names is outside the scope of `taxadb`.

@@

@@
